Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2023-12-21 doc/manuscript/translocation_pnas.tex
Minor edits to Intro and Results

Results from the population viability model were also encouraging. In
particular, translocated populations with \textgreater{} 50\% survival
in the first year post-translocation were predicted to have a low
probability of extinction over 50 years (probability of extinction \(<\)
0.5 when year 1 juvenile survival probability was greater than 0.10).
The viability model highlighted the important role of frog survival in
affecting long-term population viability, and allowed us to extend the
temporal scale of our study beyond the years covered by our
post-translocation surveys. These long-term forecasts are important,
given that reintroduced MYL frog populations may often take decades to
achieve our ultimate goal of self-sustainability \citep{joseph2018}.
Making well-supported projections about the long-term outcome of
reintroduction efforts from shorter-term information is critically
important to the process of adaptive management of species
reintroduction programs \citep{seddon2007}, including the one we are
carrying out for MYL frogs. Specifically, the combined results from our
reintroduction study and viability modeling indicate that survival of
frogs in the first year following translocation is an effective proxy of
longer-term survival and population viability. In addition, given the
repeatability of frog survival at a site, 1-year frog survival also
serves as an effective proxy of site quality (i.e., the ability of a
site to support high frog survival and a viable frog population over the
long term). This proxy of site quality is important in the MYL frog
system because accurately predicting the ability of a site to support a
viable frog population \emph{a priori} remains difficult, even after
conducting 24 translocations over 16 years.

Despite the demonstrated resistance of adult MYL frogs against Bd
infection, individual and population-level impacts of Bd are still
evident. In an earlier study of 2 of our 12 translocated populations
\citep{joseph2018}, Bd infection and load had detectable effects on the
survival of adults and may have influenced population establishment
(sites referred to as ``Alpine'' and ``Subalpine'' in \citep{joseph2018}
are identified as ``70550'' and ``70505'' in the current study).
Applying similar analyses to all 12 of our translocated populations
would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on
adults, the impacts on younger life stages are more apparent. MYL frogs
immediately following metamorphosis (``metamorphs'') are highly
susceptible to Bd infection \citep{ellison2019} and as a result
experience high mortality \citep{rachowicz2006}. This high
susceptibility of metamorphs is documented in numerous species of
anurans, and may result from the poorly developed immune system
characteristic of this life stage \citep{humphries2022}. In naturally
recovering and translocated MYL frog populations, we suggest that the
high mortality of metamorphs is an important limitation on subsequent
recruitment of new adults. Therefore, although adult MYL frogs appear
relatively resistant, Bd infection continues to have important limiting
effects on recovering populations \citep[see also][]{hollanders2022}.

The recent emergence of Bd worldwide has contributed to the decline of
hundreds of amphibian species, some of which are now extinct in the wild
\citep{scheele2019}. This extraordinary impact on global amphibian
biodiversity is compounded by the lack of any effective and broadly
applicable strategies to reverse these impacts
\citep{garner2016, knapp2022}. Importantly, in addition to the natural
recovery documented for MYL frogs \citep{knapp2016}, other species are
also showing evidence of post-epizootic recovery in the presence of Bd
\citep{scheele2017, voyles2018} and suggest the possibility of also
using animals from these recovering populations to reestablish
extirpated populations. As with MYL frogs, the feasibility and long-term
success of such efforts will depend on the availability of robust donor
populations containing individuals that have the adaptive alleles
necessary to allow frog survival and population growth in the presence
of Bd. Despite the hopeful example of successful reestablishment of MYL
frogs despite ongoing Bd infection, the challenge of recovering hundreds
of Bd-impacted amphibian species globally is a daunting prospect.
Although we now have a proven strategy to reestablish extirpated MYL
frog populations, recovery across their large historical range will
require substantial resources over many decades. The results of this
study provide a hopeful starting point for that endeavor and other
future efforts worldwide.

In our rapidly changing world, evolution is likely to play in important
role in facilitating the resilience of wildlife populations. Whether the
documented disease resistance in MYL frogs and concurrent recovery of
decimated populations provides an example of evolutionary rescue will
likely always be uncertain. Regardless, we provide an example from the
wild that suggests that evolution can produce individuals that harbor
adaptive alleles and allow population recovery in a novel (i.e.,
Bd-positive) environment, and show conclusively that individuals from
these recovering populations can be used to reestablish extirpated
populations and expand the scale of natural recovery
(Figure~\ref{fig-recovery-model}). We expect that similar species
recovery actions will be an essential tool in wildlife conservation in
an era of accelerating global change.

@@

For the 24 translocations we conducted, we identified donor populations
from which adult frogs (\(\geq\) 40 mm snout-vent length) could be
collected using several years of VES and skin swab collections
\citep{knapp2016}, and results from population genetic analyses
\citep{poorten2017}. The populations that we selected contained hundreds
to thousands of both \emph{R. sierrae} adults and tadpoles. These
relatively high abundances were the result of recent increases following
previous Bd-caused declines \citep{knapp2016}. As is typical for
recovering MYL frog populations, Bd prevalence in the donor populations
was high (0.69--0.96) and Bd load (median log\textsubscript{10}(load) =
3.06--3.78 ITS copies) was two or more orders of magnitude below the
level at which frog mortality is expected (log\textsubscript{10}(load)
\(\approx\) 5.78 copies)\citep{vredenburg2010, joseph2018}. Recipient
sites to which frogs were translocated were chosen based on previous
\emph{R. sierrae} presence (determined from VES and/or museum records)
or characteristics that suggested high quality habitat for this species
\citep{knapp2005}. At the beginning of this study, we had a relatively
limited understanding of the factors that affect habitat quality. In
subsequent years, we improved our site selection process by
incorporating new information about important habitat features, in
particular, overwinter habitats such as submerged boulders and
overhanging banks. \emph{R. sierrae} were absent from all recipient
sites prior to the first translocation.

@@

During each site visit, we conducted VES either immediately before CMR
surveys or during the first day of CMR surveys. VES was conducted by
walking the entire water body perimeter, first 100 m of each inlet and
outlet stream, and any fringing ponds and wetlands, and counting all
\emph{R. sierrae} tadpoles and juveniles. These \emph{R. sierrae} life
stages have high detectability, and counts are highly repeatable and
provide estimates of relative abundance \citep{knapp2000}.

@@

For each translocation site, we estimated survival of translocated
frogs, recruitment of new frogs into the adult population, and adult
population size using a site-specific Bayesian open-population
Jolly-Seber CMR model with known additions to the population (i.e.,
translocated cohorts), as implemented by the mrmr package
\citep{joseph2019} and using R Statistical Software
\citep[v4.4.4,][]{rsoftware2022} (see \textbf{Supporting Information -
Frog population recovery - CMR model structure} for details). Briefly,
the model tracks the states of \emph{M} individuals that comprise a
superpopulation made up of real and pseudo-individuals \citep[see][ for
details]{joseph2018}. The possible states of individuals include ``not
recruited'', ``alive'', and ``dead''. The possible observations of
individuals include ``detected'' and ``not detected''. We assume that
individuals that are in the ``not recruited'' or ``dead'' states are
never detected (i.e., there are no mistakes in the individual PIT tag
records). We also assume that new recruits were the result of
within-site reproduction and not immigration from adjacent populations.
This assumption is justified by the fact that no \emph{R. sierrae}
populations were present within several kilometers of the translocation
sites. For all models, we used mrmr defaults for priors, number of
chains (4), and warmup and post-warmup iterations (2000 for each). We
evaluated convergence of the Markov chain Monte Carlo (MCMC) algorithm
using trace plots and Gelman-Rubin statistics (Rhat).

@@

To identify important predictors of frog survival following
translocation, we used multilevel Bayesian models
\citep{gelman2013, gabry2019}. Included predictor variables describe
characteristics of sites, translocated cohorts, and individuals (Bd
load, sex, frog size, site elevation, winter severity in the year of
translocation, winter severity in the year following translocation,
donor population, day of year on which a translocation was conducted,
and translocation order). We used 1-year post-translocation survival
estimates from CMR models as the response. Estimated survival was
rounded to integer values to produce a binary outcome, and modeled with
a Bernoulli distribution. Group-level (random) effects included
site\_id, translocation\_id, or translocation\_id nested within
site\_id. We performed all analyses with the rstanarm package
\citep{rstanarm2022} and R Statistical Software
\citep[v4.4.4,][]{rsoftware2022}. For all models, we used default,
weakly informative priors, four chains, and 5000 iterations each for
warmup and post-warmup. We checked MCMC convergence using trace plots
and Rhat, and evaluated model fit using leave-one-out cross-validation
\citep{vehtari2016}, as implemented by the loo package
\citep{vehtari2022}. (See \textbf{Supporting Information - Frog
population recovery - Among-site survival modeling} for details.)

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods
\citep[see \textbf{Supporting Information-Frog population recovery -
Laboratory methods} for details]{boyle2004}. To assess the magnitude of
changes in Bd load on frogs following translocation, we compared Bd
loads measured before versus after translocation. Before-translocation
loads were quantified using skin swabs collected from all
to-be-translocated frogs at the donor site on the day before or the day
of the translocation. After-translocation Bd loads were based on all
swabs collected from translocated frogs at the recipient site in the
year of and the year following translocation. Individual frogs and their
associated Bd loads were included in the dataset only if frogs were
captured at the recipient site at least once during the 1-year period
following translocation.

@@

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence \citep{joseph2018}, and (ii) host density does not seem
to play a significant role in multi-year Bd infection dynamics in this
system \citep{wilber2022}. Thus, ignoring Bd infection dynamics and
instead assuming all host vital rates are in the presence of high Bd
prevalence significantly simplifies the model without much loss of
realism. Additional details are provided in \textbf{Supporting
Information - Population viability modeling}.

@@

We collected DNA samples via buccal swabbing \citep{broquet2007} from 53
\emph{Rana muscosa}/\emph{Rana sierrae} individuals: 24 from 4 naive
populations, and 29 from 5 recovering populations. These populations are
located in the southern Sierra Nevada, from northern Yosemite National
Park to northern Sequoia National Park
(Figure~\ref{fig-allelefrequencies}). Samples were collected from 5-6
frogs per population. To minimize potential confounding effects caused
by variation in frog genotypes across latitude \citep{byrne2023}, we
selected sampling sites such that both population types were represented
across similar latitudinal ranges. DNA was extracted following Qiagen
DNEasy manufacturer's protocols. We sequenced the samples using an exome
capture assay as described in \citep{byrne2023}. Briefly, genomic
libraries were prepared and captured using a custom Nimblegen capture
pool. Capture baits were designed based on the coding regions of the
\emph{R. muscosa} transcriptome (GenBank accession GKCT00000000).
Captured libraries were then pooled and sequenced on a NovaSeq 6000
150PE Flow Cell S1 at the Vincent J. Coates Genomics Sequencing Lab at
UC Berkeley. Raw sequencing reads are available from NCBI SRA
(PRJNA870451).

@@

Raw reads were filtered for adapters and contaminants using fastp
\citep{chen2018} and aligned to the \emph{Rana muscosa} genome (NCBI
SRA: SRS6533475 \citep{hon2020}), with repetitive elements masked using
bwa (``mem'' mode) \citep{li2013}. Exact PCR duplicates were marked
using Picard. Variants were then called following GATK best practices
(v.4.2.0.0 \citep{van2020}). Briefly, raw variants were called for each
sample using HaplotypeCaller and combined using CombineGVCFs. Next,
genotypes were jointly called using GenotypeGVCFs. Variants were then
hard filtered using gatk VariantFiltration using the following
parameters to remove low-quality sites: QD \textless{} 2.0, FS
\textgreater{} 40.0, SOR \textgreater{} 3.0, MQ \textless{} 50.0,
MQRankSum \textgreater{} 3.0, MQRankSum \textless{} -3.0, ReadPosRankSum
\textgreater{} 3.0, ReadPosRankSum \textless{} -3.0. This initial filter
resulted in 1,595,206 variant sites across 53 individuals.

@@

To visualize the genomic relationships of our populations, we conducted
a PCA using the glPCA function in the adegenet R package
\citep{jombart2008}. To detect regions of the genome that differed
between naive and recovering populations, i.e., regions putatively under
selection, we used two approaches: (1) a multivariate linear mixed model
to evaluate individual variants (SNPs and INDELs), and (2) a splined
window analysis to evaluate larger genomic regions. For the variant
analysis, we first used a stringent data filter to include only variants
with \textless{} 5\% missing data (missing for no more than 2
individuals), and then calculated the likelihood ratio statistic for the
resulting set of 148,307 high quality variants across 127 contigs using
GEMMA \citep{zhou2014}. GEMMA calculates and incorporates a relatedness
matrix for input samples, allowing us to account for relatedness and
population structure when calculating likelihood ratio statistics. We
identified variants showing different allele frequencies between naive
versus recovering populations (``outliers'') using a
Bonferroni-corrected significance level of 0.01. We visualized the
results using a Manhattan plot and qqplot. We developed a more liberal
set of outlier variants using a Bonferroni-corrected significance level
of 0.05 and used this set solely for the gene ontology (GO) analysis
(see below and \textbf{Supporting Information}: Dataset S1, S2).

In the analysis of individual variants, for each outlier variant we
determined whether the variant was synonymous (protein sequence the same
for each variant) or non-synonymous (protein sequence differs between
variants), and where in the gene it was located. To do this, we first
extracted the reference genome sequence surrounding the variant using
the bedtools ``getfasta'' function \citep{quinlan2010}. Next, we
re-annotated each sequence using BLAST to get the predicted gene
location based on the closest annotated reference \citep{altschul1997}.
We then translated each variant to amino acids and aligned this
translation to that of the gene annotation to ensure proper frame of
reference using Geneious \citep{kearse2012}. After ensuring proper
translation, we characterized variants as within or outside the coding
sequence of the gene and as either synonymous or non-synonymous.

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools \citep{danecek2011}. Next, we
calculated per-site nucleotide diversity \(\pi\) separately for
individuals from the naive and recovering populations using VCFtools,
then calculated \(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome (for more efficient data processing and
to better contextualize the strength of the outlier signals in different
regions of the genome). We then used the GenWin R package
\citep{beissinger2015} to conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the spline fitted to \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) values along the genome, therefore balancing false
positive and false negative results that occur using other window-based
methods \citep{beissinger2015}. This method also calculates a
W-statistic allowing for outlier identification. We identified outliers
as those with a W-statistic greater than 4 standard deviations above the
mean for \emph{F\textsubscript{ST}} or above/below the mean for
\(\pi_{diff}\). These standard deviations represent strict criteria to
select only the top \textasciitilde{} 0.3\% of windows. Shared outliers
were then identified as those that were outliers in both analyses,
meaning that they showed (i) high differentiation between naive and
recovering populations, and (ii) differential patterns of nucleotide
diversity in the same region. Finally, we extracted gene transcripts
mapped within each region and retrieved annotation for that region using
BLAST (\textbf{Supporting Information}: Datasets S3, S4).

@@

\caption{\label{fig-recovery-model}For \emph{R. sierrae}, a conceptual
model depicting the Bd-caused decline and subsequent natural recovery
(black text), facilitated recovery via reintroductions, and the linkages
between these two pathways. Rectangles and hexagons represent outcomes
and processes, respectively. Blue text indicates components that are
included in the current study. The timeline shows the general sequence
of the components, with the dotted portion indicating a projection into
the future.}

@@

\caption{\label{fig-cond-effects}Results show conditional effects of the
important predictors of 1-year frog survival (survival is expressed as a
probability), as estimated from the model used in the among-site
meta-analysis: (A) winter severity in the year following translocation,
(B) elevation of recipient site, (C) donor population, and (D) sex. In
(A) and (B), blue lines are medians and gray ribbons are 95\%
uncertainty intervals. In (C) and (D), box plots show medians, first and
third quartiles, largest and smallest values within 1.5x interquartile
range, and values outside the 1.5x interquartile range.}

>>>>

Results from the population viability model were also encouraging. In
particular, translocated populations with \textgreater{} 50\% survival
in the first year post-translocation were predicted to have a low
probability of extinction over 50 years (probability of extinction \(<\)
0.5 when year 1 juvenile survival probability was greater than 0.10).
The viability model highlighted the important role of frog survival in
affecting long-term population viability, and allowed us to extend the
temporal scale of our study beyond the years covered by our
post-translocation surveys. These long-term forecasts are important,
given that reintroduced MYL frog populations may often take decades to
achieve our ultimate goal of self-sustainability (41). Making
well-supported projections about the long-term outcome of reintroduction
efforts from shorter-term information is critically important to the
process of adaptive management of species reintroduction programs (62),
including the one we are carrying out for MYL frogs. Specifically, the
combined results from our reintroduction study and viability modeling
indicate that survival of frogs in the first year following
translocation is an effective proxy of longer-term survival and
population viability. In addition, given the repeatability of frog
survival at a site, 1-year frog survival also serves as an effective
proxy of site quality (i.e., the ability of a site to support high frog
survival and a viable frog population over the long term). This proxy of
site quality is important in the MYL frog system because accurately
predicting the ability of a site to support a viable frog population
\emph{a priori} remains difficult, even after conducting 24
translocations over 16 years.

Despite the demonstrated resistance of adult MYL frogs against Bd
infection, individual and population-level impacts of Bd are still
evident. In an earlier study of 2 of our 12 translocated populations
(41), Bd infection and load had detectable effects on the survival of
adults and may have influenced population establishment (sites referred
to as ``Alpine'' and ``Subalpine'' in (41) are identified as ``70550''
and ``70505'' in the current study). Applying similar analyses to all 12
of our translocated populations would likely provide a broader
perspective of the ongoing effect of Bd. In addition to these important
but relatively subtle effects of Bd on adults, the impacts on younger
life stages are more apparent. MYL frogs immediately following
metamorphosis (``metamorphs'') are highly susceptible to Bd infection
(63) and as a result experience high mortality (34). This high
susceptibility of metamorphs is documented in numerous species of
anurans, and may result from the poorly developed immune system
characteristic of this life stage (64). In naturally recovering and
translocated MYL frog populations, we suggest that the high mortality of
metamorphs is an important limitation on subsequent recruitment of new
adults. Therefore, although adult MYL frogs appear relatively resistant,
Bd infection continues to have important limiting effects on recovering
populations (see also 65).

The recent emergence of Bd worldwide has contributed to the decline of
hundreds of amphibian species, some of which are now extinct in the wild
(8). This extraordinary impact on global amphibian biodiversity is
compounded by the lack of any effective and broadly applicable
strategies to reverse these impacts (17, 25). Importantly, in addition
to the natural recovery documented for MYL frogs (14), other species are
also showing evidence of post-epizootic recovery in the presence of Bd
(12, 13) and suggest the possibility of also using animals from these
recovering populations to reestablish extirpated populations. As with
MYL frogs, the feasibility and long-term success of such efforts will
depend on the availability of robust donor populations containing
individuals that have the adaptive alleles necessary to allow frog
survival and population growth in the presence of Bd. Despite the
hopeful example of successful reestablishment of MYL frogs despite
ongoing Bd infection, the challenge of recovering hundreds of
Bd-impacted amphibian species globally is a daunting prospect. Although
we now have a proven strategy to reestablish extirpated MYL frog
populations, recovery across their large historical range will require
substantial resources over many decades. The results of this study
provide a hopeful starting point for that endeavor and other future
efforts worldwide.

In our rapidly changing world, evolution is likely to play in important
role in facilitating the resilience of wildlife populations. Whether the
documented disease resistance in MYL frogs and concurrent recovery of
decimated populations provides an airtight example of evolutionary
rescue will likely always be uncertain (given that can never have a
perfect understanding of the past). Regardless, we provide an example
from the wild that suggests that evolution can produce individuals that
harbor adaptive alleles and allow population recovery in a novel (i.e.,
Bd-positive) environment, and show conclusively that individuals from
these recovering populations can be used to reestablish extirpated
populations and expand the scale of natural recovery
(Figure~\ref{fig-recovery-model}). We expect that similar species
recovery actions will be an essential tool in wildlife conservation in
an era of accelerating global change.

@@

For the 24 translocations we conducted, we identified donor populations
from which adult frogs (\(\geq\) 40 mm snout-vent length) could be
collected using several years of VES and skin swab collections (14), and
results from population genetic analyses (48). The populations that we
selected contained hundreds to thousands of both \emph{R. sierrae}
adults and tadpoles. These relatively high abundances were the result of
recent increases following previous Bd-caused declines (14). As is
typical for recovering MYL frog populations, Bd prevalence in the donor
populations was high (0.69--0.96) and Bd load (median
log\textsubscript{10}(load) = 3.06--3.78 ITS copies) was two or more
orders of magnitude below the level at which frog mortality is expected
(log\textsubscript{10}(load) \(\approx\) 5.78 copies)(33, 41). Recipient
sites to which frogs were translocated were chosen based on previous
\emph{R. sierrae} presence (determined from VES and/or museum records)
or characteristics that suggested high quality habitat for this species
(66). At the beginning of this study, we had a relatively limited
understanding of the factors that affect habitat quality. In subsequent
years, we improved our site selection process by incorporating new
information about important habitat features, in particular, overwinter
habitats such as submerged boulders and overhanging banks. \emph{R.
sierrae} were absent from all recipient sites prior to the first
translocation.

@@

During each site visit, we conducted VES either immediately before CMR
surveys or during the first day of CMR surveys. VES was conducted by
walking the entire water body perimeter, first 100 m of each inlet and
outlet stream, and any fringing ponds and wetlands, and counting all
\emph{R. sierrae} tadpoles and juveniles. These \emph{R. sierrae} life
stages have high detectability, and counts are highly repeatable and
provide estimates of relative abundance (31).

@@

For each translocation site, we estimated survival of translocated
frogs, recruitment of new frogs into the adult population, and adult
population size using a site-specific Bayesian open-population
Jolly-Seber CMR model with known additions to the population (i.e.,
translocated cohorts), as implemented by the mrmr package (67) and using
R Statistical Software (v4.4.4, 68) (see \textbf{Supporting Information
- Frog population recovery - CMR model structure} for details). Briefly,
the model tracks the states of \emph{M} individuals that comprise a
superpopulation made up of real and pseudo-individuals (see 41 for
details). The possible states of individuals include ``not recruited'',
``alive'', and ``dead''. The possible observations of individuals
include ``detected'' and ``not detected''. We assume that individuals
that are in the ``not recruited'' or ``dead'' states are never detected
(i.e., there are no mistakes in the individual PIT tag records). We also
assume that new recruits were the result of within-site reproduction and
not immigration from adjacent populations. This assumption is justified
by the fact that no \emph{R. sierrae} populations were present within
several kilometers of the translocation sites. For all models, we used
mrmr defaults for priors, number of chains (4), and warmup and
post-warmup iterations (2000 for each). We evaluated convergence of the
Markov chain Monte Carlo (MCMC) algorithm using trace plots and
Gelman-Rubin statistics (Rhat).

@@

To identify important predictors of frog survival following
translocation, we used multilevel Bayesian models (69, 70). Included
predictor variables describe characteristics of sites, translocated
cohorts, and individuals (Bd load, sex, frog size, site elevation,
winter severity in the year of translocation, winter severity in the
year following translocation, donor population, day of year on which a
translocation was conducted, and translocation order). We used 1-year
post-translocation survival estimates from CMR models as the response.
Estimated survival was rounded to integer values to produce a binary
outcome, and modeled with a Bernoulli distribution. Group-level (random)
effects included site\_id, translocation\_id, or translocation\_id
nested within site\_id. We performed all analyses with the rstanarm
package (71) and R Statistical Software (v4.4.4, 68). For all models, we
used default, weakly informative priors, four chains, and 5000
iterations each for warmup and post-warmup. We checked MCMC convergence
using trace plots and Rhat, and evaluated model fit using leave-one-out
cross-validation (72), as implemented by the loo package (73). (See
\textbf{Supporting Information - Frog population recovery - Among-site
survival modeling} for details.)

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods
(74, see \textbf{Supporting Information-Frog population recovery -
Laboratory methods} for details). To assess the magnitude of changes in
Bd load on frogs following translocation, we compared Bd loads measured
before versus after translocation. Before-translocation loads were
quantified using skin swabs collected from all to-be-translocated frogs
at the donor site on the day before or the day of the translocation.
After-translocation Bd loads were based on all swabs collected from
translocated frogs at the recipient site in the year of and the year
following translocation. Individual frogs and their associated Bd loads
were included in the dataset only if frogs were captured at the
recipient site at least once during the 1-year period following
translocation.

@@

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence (41), and (ii) host density does not seem to play a
significant role in multi-year Bd infection dynamics in this system
(46). Thus, ignoring Bd infection dynamics and instead assuming all host
vital rates are in the presence of high Bd prevalence significantly
simplifies the model without much loss of realism. Additional details
are provided in \textbf{Supporting Information - Population viability
modeling}.

@@

We collected DNA samples via buccal swabbing (75) from 53 \emph{Rana
muscosa}/\emph{Rana sierrae} individuals: 24 from 4 naive populations,
and 29 from 5 recovering populations. These populations are located in
the southern Sierra Nevada, from northern Yosemite National Park to
northern Sequoia National Park (Figure~\ref{fig-allelefrequencies}).
Samples were collected from 5-6 frogs per population. To minimize
potential confounding effects caused by variation in frog genotypes
across latitude (49), we selected sampling sites such that both
population types were represented across similar latitudinal ranges. DNA
was extracted following Qiagen DNEasy manufacturer's protocols. We
sequenced the samples using an exome capture assay as described in (49).
Briefly, genomic libraries were prepared and captured using a custom
Nimblegen capture pool. Capture baits were designed based on the coding
regions of the \emph{R. muscosa} transcriptome (GenBank accession
GKCT00000000). Captured libraries were then pooled and sequenced on a
NovaSeq 6000 150PE Flow Cell S1 at the Vincent J. Coates Genomics
Sequencing Lab at UC Berkeley. Raw sequencing reads are available from
NCBI SRA (PRJNA870451).

@@

Raw reads were filtered for adapters and contaminants using fastp (76)
and aligned to the \emph{Rana muscosa} genome (NCBI SRA: SRS6533475
(77)), with repetitive elements masked using bwa (``mem'' mode) (78).
Exact PCR duplicates were marked using Picard. Variants were then called
following GATK best practices (v.4.2.0.0 (79)). Briefly, raw variants
were called for each sample using HaplotypeCaller and combined using
CombineGVCFs. Next, genotypes were jointly called using GenotypeGVCFs.
Variants were then hard filtered using gatk VariantFiltration using the
following parameters to remove low-quality sites: QD \textless{} 2.0, FS
\textgreater{} 40.0, SOR \textgreater{} 3.0, MQ \textless{} 50.0,
MQRankSum \textgreater{} 3.0, MQRankSum \textless{} -3.0, ReadPosRankSum
\textgreater{} 3.0, ReadPosRankSum \textless{} -3.0. This initial filter
resulted in 1,595,206 variant sites across 53 individuals.

@@

To visualize the genomic relationships of our populations, we conducted
a PCA using the glPCA function in the adegenet R package (80). To detect
regions of the genome that differed between naive and recovering
populations, i.e., regions under selection, we used two approaches: (1)
a multivariate linear mixed model to evaluate individual variants (SNPs
and INDELs), and (2) a splined window analysis to evaluate larger
genomic regions. For the variant analysis, we first used a stringent
data filter to include only variants with \textless{} 5\% missing data
(missing for no more than 2 individuals), and then calculated the
likelihood ratio statistic for the resulting set of 148,307 high quality
variants across 127 contigs using GEMMA (81). GEMMA calculates and
incorporates a relatedness matrix for input samples, allowing us to
account for relatedness and population structure when calculating
likelihood ratio statistics. We identified variants showing different
allele frequencies between naive versus recovering populations
(``outliers'') using a Bonferroni-corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot. We
developed a more liberal set of outlier variants using a
Bonferroni-corrected significance level of 0.05 and used this set solely
for the gene ontology (GO) analysis (see below and \textbf{Supporting
Information}: Dataset S1, S2).

In the analysis of individual variants, for each outlier variant we
determined whether the variant was synonymous (protein sequence the same
for each variant) or non-synonymous (protein sequence differs between
variants), and where in the gene it was located. To do this, we first
extracted the reference genome sequence surrounding the variant using
the bedtools ``getfasta'' function (82). Next, we re-annotated each
sequence using BLAST to get the predicted gene location based on the
closest annotated reference (83). We then translated each variant to
amino acids and aligned this translation to that of the gene annotation
to ensure proper frame of reference using Geneious (84). After ensuring
proper translation, we characterized variants as within or outside the
coding sequence of the gene and as either synonymous or non-synonymous.

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools (85). Next, we calculated per-site
nucleotide diversity \(\pi\) separately for individuals from the naive
and recovering populations using VCFtools, then calculated
\(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome (for more efficient data processing and
to better contextualize the strength of the outlier signals in different
regions of the genome). We then used the GenWin R package (86) to
conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the spline fitted to \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) values along the genome, therefore balancing false
positive and false negative results that occur using other window-based
methods (86). This method also calculates a W-statistic allowing for
outlier identification. We identified outliers as those with a
W-statistic greater than 4 standard deviations above the mean for
\emph{F\textsubscript{ST}} or above/below the mean for \(\pi_{diff}\).
These standard deviations represent strict criteria to select only the
top \textasciitilde{} 0.3\% of windows. Shared outliers were then
identified as those that were outliers in both analyses, meaning that
they showed (i) high differentiation between naive and recovering
populations, and (ii) differential patterns of nucleotide diversity in
the same region. Finally, we extracted gene transcripts mapped within
each region and retrieved annotation for that region using BLAST
(\textbf{Supporting Information}: Datasets S3, S4).

@@

\caption{\label{fig-recovery-model}For MYL frogs, a conceptual model
depicting the Bd-caused decline and subsequent natural recovery (black
text), facilitated recovery via reintroductions, and the linkages
between these two pathways. Rectangles and hexagons represent outcomes
and processes, respectively. Blue text indicates components that are
included in the current study. The timeline shows the general sequence
of the components, with the dotted portion indicating a projection into
the future.}

@@

\caption{\label{fig-cond-effects}Results from the among-site
meta-analysis showing conditional effects of the important predictors of
1-year frog survival (expressed as a probability): (A) winter severity
in the year following translocation, (B) elevation of recipient site,
(C) donor population, and (D) sex. In (A) and (B), blue lines are
medians and gray ribbons are 95\% uncertainty intervals. In (C) and (D),
box plots show medians, first and third quartiles, largest and smallest
values within 1.5x interquartile range, and values outside the 1.5x
interquartile range.}
